---
title: "Data Exploration"
author: "Peter Madsen and Chris Salas"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(readr)
library(dplyr)
library(forcats)

# TODO: Consider using source() for script files to make things tidy.
# Also, here's how to save graphs https://ggplot2.tidyverse.org/reference/ggsave.html

fullsomerville2017 <- read_csv("cleanedData/2017HappinessSurvey.csv")

```

# Exploration

The 2017 survey data was cleaned up in Python (see `data_cleanup_changelog.md`) and exported into the `cleanedData` folder. 

## Goals (Nov/11/2020)

> 1. Find Somerville population demographics to compare our sample demographics to
  - Useful starting point: https://www.census.gov/quickfacts/fact/table/somervillecitymassachusetts/PST045219
  
### Removing blanks and low-n race/ethnicity

``` {r removing_blanks}

# blanks
cat("Before filter:", nrow(fullsomerville2017), "\n")  # 835
fullsomerville2017 = fullsomerville2017 %>% filter(`2. Age?` != "Blank")
cat("After removing age-blank:", nrow(fullsomerville2017), "\n")  # 821 (-14)
fullsomerville2017 = fullsomerville2017 %>% filter(`9. What is your annual household income?` != "Blank")
cat("After removing income-blank:", nrow(fullsomerville2017), "\n")  # 736 (-85)
fullsomerville2017 = fullsomerville2017 %>% filter(`4. What is your race or ethnicity?` != "Blank")
cat("After removing race/ethnicity-blank:", nrow(fullsomerville2017), "\n")  # 709 (-27)
fullsomerville2017 = fullsomerville2017 %>% filter(`1. What is your gender` != "Blank")
cat("After removing gender-blank: ", nrow(fullsomerville2017), "\n")  # 703 (-6)

# low-n race/ethnicity
fullsomerville2017 = fullsomerville2017 %>% filter(`4. What is your race or ethnicity?` == "White" | `4. What is your race or ethnicity?` == "Asian" | `4. What is your race or ethnicity?` == "Hispanic/Latino" | `4. What is your race or ethnicity?` == "Black/African American")
cat("After removing low-n race/ethnicity:", nrow(fullsomerville2017), "\n")  # 662 (-41)

# Removing blanks for house price and night safety
fullsomerville2017 = fullsomerville2017 %>% filter(`7b. How would you rate the cost of housing?` != 0)
cat("After removing house price zeros:", nrow(fullsomerville2017), "\n")  # 653 (-9)
fullsomerville2017 = fullsomerville2017 %>% filter(`9. How safe do you feel walking in your neighborhood at night?` != 0)
cat("After removing night safety zeros:", nrow(fullsomerville2017), "\n")  # 647 (-6)

# Removing blanks for police trust
fullsomerville2017 = fullsomerville2017 %>% filter(`7d. How would you rate your trust in the local police?` != 0)
cat("After removing police trust zeros:", nrow(fullsomerville2017), "\n")  # 631 (-16)

cat("Total removed: ", (835 - 631), " (", (100 * (835-631) / 835), "%)\n", sep = "")

```



### Frequency bar plots


```{r demographics_barplots}
# Demographics in a single dataframe.
demographicssomerville2017 <- fullsomerville2017[c(18:23, 25:27)]


# Each column of interest in a single dataframe. (So we can see)
# sample distribution of demographics)
gender = demographicssomerville2017[1]
age = demographicssomerville2017[2]
ethnicity = demographicssomerville2017[4]
annual_income = demographicssomerville2017[8]
housing_status = demographicssomerville2017[6]
student = demographicssomerville2017[9]

renters = fullsomerville2017 %>% filter(`6. Describe your housing status in Somerville.` == "Rent")
owners = fullsomerville2017 %>% filter(`6. Describe your housing status in Somerville.` == "Own")

house_price = fullsomerville2017[c(9)]


# Counting up unique entries
# gender %>% count(gender[1])
# age %>% count(age[1])
# ethnicity %>% count(ethnicity[1])
# annual_income %>% count(annual_income[1])
# renter %>% count(renter[1])
# student %>% count(student[1])


# Choosing the order of the x-axis entries
gender_ordered = gender %>% dplyr::mutate(`1. What is your gender` = factor(`1. What is your gender`, levels=c("Blank", "Female", "Male", "Non-binary", "Both selected", "Multiple people")))
age_ordered = age %>% dplyr::mutate(`2. Age?` = factor(`2. Age?`, levels=c("Blank", "18-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75 or older")))
annual_income_ordered = annual_income %>% dplyr::mutate(`9. What is your annual household income?` = factor(`9. What is your annual household income?`, levels=c("Blank", "Less than $10,000", "$10,000 to $24,999", "$25,000 to $49,999", "$50,000 to 74,999", "$75,000 to $99,999", "$100,000 to $149,999", "$150,000 or more")))


#The bar plots of each.
ggplot(gender_ordered, aes_string(x = "`1. What is your gender`")) + geom_bar() + labs(x = "", y = "", title = "\"What is your gender?\"")

ggplot(age_ordered, aes(x = `2. Age?`)) + geom_bar() + labs(x = "", y = "", title = "\"What is your age?\"")

ggplot(ethnicity, aes_string(x = "`4. What is your race or ethnicity?`")) + geom_bar() + labs(x = "", y = "", title = "\"What is your race or ethnicity?\"") + theme(axis.text.x = element_text(angle = 89)) + theme(axis.text.x = element_text(vjust = 0.5)) + coord_flip() 

ggplot(annual_income_ordered, aes_string(x = "`9. What is your annual household income?`")) + geom_bar() + labs(x = "", y = "", title = "\"What is your annual household income?\"") + theme(axis.text.x = element_text(angle = 75)) + theme(axis.text.x = element_text(vjust = 0.5)) 

ggplot(housing_status, aes(x = `6. Describe your housing status in Somerville.`)) + geom_bar() + labs(x = "", y = "", title = "\"Describe your housing status\"")

ggplot(student, aes(x = `10. Are you a student?`)) + geom_bar() + labs(x = "", y = "", title = "\"Are you a student?\"")

ggplot(house_price, aes(x = `7b. How would you rate the cost of housing?`)) + geom_bar() + labs(x = "", y = "", title = "\"How would you rate the cost of housing?\"") + scale_x_continuous(n.breaks = 6)


```

### Functions for stacked bar and separated distributions

``` {r the_functions}

# ----- Begin stacked bar function definition.
make_stacked_bars = function(column, xticks, dataframe=fullsomerville2017) {

gender_stacked = dataframe %>% 
  dplyr::mutate(`1. What is your gender` = factor(`1. What is your gender`, levels=c("Blank", "Female", "Male", "Non-binary", "Both selected", "Multiple people")))
w = dataframe %>% 
  ggplot(aes_string(x=column, fill="`1. What is your gender`")) + geom_bar() + labs(x = "", y = "", title = column) + scale_x_continuous(n.breaks = xticks)
print(w)

age_stacked = dataframe %>% dplyr::mutate(`2. Age?` = factor(`2. Age?`, levels=c("Blank", "18-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75 or older")))
x = ggplot(age_stacked, aes_string(x=column, fill="`2. Age?`")) + geom_bar() + labs(x = "", y = "", title = column) + scale_x_continuous(n.breaks = xticks)
print(x)

ethnicity_stacked = dataframe %>% filter(`4. What is your race or ethnicity?` == "White" | `4. What is your race or ethnicity?` == "Blank" | `4. What is your race or ethnicity?` == "Asian" | `4. What is your race or ethnicity?` == "Hispanic/Latino" | `4. What is your race or ethnicity?` == "Black/African American")
y = ethnicity_stacked %>% ggplot(aes_string(x=column, fill="`4. What is your race or ethnicity?`")) + geom_bar() + labs(x = "", y = "", title = column) + scale_x_continuous(n.breaks = xticks)
print(y)

annual_income_stacked = dataframe %>% dplyr::mutate(`9. What is your annual household income?` = factor(`9. What is your annual household income?`, levels=c("Blank", "Less than $10,000", "$10,000 to $24,999", "$25,000 to $49,999", "$50,000 to 74,999", "$75,000 to $99,999", "$100,000 to $149,999", "$150,000 or more")))
z = annual_income_stacked %>% ggplot(aes_string(x=column, fill="`9. What is your annual household income?`")) + geom_bar() + labs(x = "", y = "", title = column) + scale_x_continuous(n.breaks = xticks)
print(z)
}  # end stacked bar function definition


# ----- begin demographic subset function definition
demographic_subset_charts = function(column, xticks, dataframe=fullsomerville2017) {
  gender_subsets = c("Female", 
                     "Male", 
                     "Non-binary", 
                     "Both selected", 
                     "Multiple people")
  age_subsets = c("18-24", 
                  "25-34", 
                  "35-44", 
                  "45-54", 
                  "55-64", 
                  "65-74", 
                  "75 or older")
  income_subsets = c("Less than $10,000", 
                     "$10,000 to $24,999", 
                     "$25,000 to $49,999", 
                     "$50,000 to 74,999", 
                     "$75,000 to $99,999", 
                     "$100,000 to $149,999", 
                     "$150,000 or more")
  ethnicity_subsets = c("White", 
                        "Asian", 
                        "Hispanic/Latino", 
                        "Black/African American")
  student_subsets = c("Blank", 
                      "False", 
                      "True")
  renter_subsets = c("Neither", 
                     "Own", 
                     "Rent")
  
  gender_data = dataframe %>% dplyr::mutate(`1. What is your gender` = factor(`1. What is your gender`, levels=c("Female", "Male", "Non-binary", "Both selected", "Multiple people")))
  age_data = dataframe %>% dplyr::mutate(`2. Age?` = factor(`2. Age?`, levels=c("18-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75 or older")))
  income_data = dataframe %>% dplyr::mutate(`9. What is your annual household income?` = factor(`9. What is your annual household income?`, levels=c("Less than $10,000", "$10,000 to $24,999", "$25,000 to $49,999", "$50,000 to 74,999", "$75,000 to $99,999", "$100,000 to $149,999", "$150,000 or more")))
  ethnicity_data = dataframe 
  student_data = dataframe
  renter_data = dataframe


for (i in gender_subsets) {
  w = gender_data %>% 
    filter(`1. What is your gender` == i) %>% 
    ggplot(aes_string(x=column, fill="`1. What is your gender`")) + geom_bar(color = "black") + labs(x = "", y = "", title = column) + scale_x_continuous(breaks = c(1:xticks), limits = c(0, (xticks+1))) + scale_fill_manual(values = c("red"))
  print(w)
}
for (i in age_subsets) {
  w = age_data %>% 
    filter(`2. Age?` == i) %>% 
    ggplot(aes_string(x=column, fill="`2. Age?`")) + geom_bar(color = "black") + labs(x = "", y = "", title = column) + scale_x_continuous(breaks = c(1:xticks), limits = c(0, (xticks+1))) + scale_fill_manual(values = c("blue"))
  print(w)
}
for (i in income_subsets) {
  w = income_data %>% 
    filter(`9. What is your annual household income?` == i) %>%
    ggplot(aes_string(x=column, fill="`9. What is your annual household income?`")) + geom_bar(color = "black") + labs(x = "", y = "", title = column) + scale_x_continuous(breaks = c(1:xticks), limits = c(0, (xticks+1))) + scale_fill_manual(values = c("green"))
  print(w)
}
for (i in ethnicity_subsets) {
  w = ethnicity_data %>% 
    filter(`4. What is your race or ethnicity?` == i) %>% 
    ggplot(aes_string(x=column, fill="`4. What is your race or ethnicity?`")) + geom_bar(color = "black") + labs(x = "", y = "", title = column) + scale_x_continuous(breaks = c(1:xticks), limits = c(0, (xticks+1))) + scale_fill_manual(values = c("purple"))
  print(w)
}
for (i in student_subsets) {
  w = student_data %>% 
    filter(`10. Are you a student?` == i) %>% 
    ggplot(aes_string(x=column, fill="`10. Are you a student?`")) + geom_bar(color = "black") + labs(x = "", y = "", title = column) + scale_x_continuous(breaks = c(1:xticks), limits = c(0, (xticks+1))) + scale_fill_manual(values = c("orange"))
  print(w)
}
for (i in renter_subsets) {
  w = renter_data %>% 
    filter(`6. Describe your housing status in Somerville.` == i) %>% 
    ggplot(aes_string(x=column, fill="`6. Describe your housing status in Somerville.`")) + geom_bar(color = "black") + labs(x = "", y = "", title = column) + scale_x_continuous(breaks = c(1:xticks), limits = c(0, (xticks+1))) + scale_fill_manual(values = c("cyan"))
  print(w)
}
  
}  # end demographic subset function definition


# Begin renter/owner function
rent_own = function(column, xticks, renter = renters, owner = owners, categorical = FALSE) {
  if (categorical == FALSE) {
    r = renters %>% ggplot(aes_string(x=column, fill="`6. Describe your housing status in Somerville.`")) + geom_bar(color = "black") + labs(x = "", y = "", title = column) + scale_x_continuous(breaks = c(1:xticks), limits = c(0, (xticks+1))) + scale_fill_manual(values = c("royalblue3"))
    print(r)
  
    o = owners %>% ggplot(aes_string(x=column, fill="`6. Describe your housing status in Somerville.`")) + geom_bar(color = "black") + labs(x = "", y = "", title = column) + scale_x_continuous(breaks = c(1:xticks), limits = c(0, (xticks+1))) + scale_fill_manual(values = c("springgreen4"))
    print(o)
  }
  else {
    r = renters %>% ggplot(aes_string(x=column, fill="`6. Describe your housing status in Somerville.`")) + geom_bar(color = "black") + labs(x = "", y = "", title = column) + scale_fill_manual(values = c("steelblue1"))
    print(r)
  
    o = owners %>% ggplot(aes_string(x=column, fill="`6. Describe your housing status in Somerville.`")) + geom_bar(color = "black") + labs(x = "", y = "", title = column) + scale_fill_manual(values = c("seagreen1"))
    print(o)
  }
}


```



### Demographics vs. House Prices



``` {r houseprice_bars}

# Behold! All these charts with a single line!
make_stacked_bars(column = "`7b. How would you rate the cost of housing?`", xticks = 6)

demographic_subset_charts(column = "`7b. How would you rate the cost of housing?`", xticks = 5)
# TODO: Verify what the numbers in the hedonic scale represent
```

### Demographics vs. Night Safety

``` {r demographics_night}
make_stacked_bars(column = "`9. How safe do you feel walking in your neighborhood at night?`", xticks = 11)

demographic_subset_charts(column = "`9. How safe do you feel walking in your neighborhood at night?`", xticks = 10)

# TODO: Box plots of these
# TODO: Hypothesis testing (see lab 5.3)

```

### Demographics vs. Police Trust

``` {r demographics_police}
make_stacked_bars(column = "`7d. How would you rate your trust in the local police?`", xticks = 6)

demographic_subset_charts(column = "`7d. How would you rate your trust in the local police?`", xticks = 5)
```

### Demographics vs. EVERYTHING!

``` {r demographics_all}

# This is a list of all the column headers with "rate this on a scale
# from 1 to n" kinda questions. Putting them all here lets me loop
# through them.
all_rating_questions = c("`1. How happy do you feel right now?`", 
                         "`2. How satisfied are you with your life in general?`", 
                         "`3. How anxious did you feel yesterday?`", 
                         "`4. How satisfied are you with Somerville as a place to live?`", 
                         "`5. How satisfied are you with your neighborhood?`", 
                         "`6. How proud are you to be a Somerville resident?`", 
                         "`7a. How would you rate the availability of information about city services?`", 
                         "`7b. How would you rate the cost of housing?`", 
                         "`7c. How would you rate the overall quality of public schools?`", 
                         "`7d. How would you rate your trust in the local police?`", 
                         "`7e. How would you rate the maintenance of streets and sidewalks?`", 
                         "`7f. How would you rate the availability of social community events?`", 
                         "`8. How satisfied are you with the quality and number of transportation options available to you?`", 
                         "`9. How safe do you feel walking in your neighborhood at night?`", 
                         "`10. How satisfied are you with the appearance of parks and squares in your neighborhood?`", 
                         "`11. How satisfied are you with the beauty or physical setting of your neighborhood?`")

all_categorical_questions = c("`1. What is your gender`", 
                              "`2. Age?`", 
                              "`3. What language(s) do you speak at home?`", 
                              "`4. What is your race or ethnicity?`", 
                              "`5. Do you have children age 18 or younger who live with you?`", 
                              "`6. Describe your housing status in Somerville.`", 
                              "`7. Do you plan to move away from Somerville in the next two years?`", 
                              "`8 How long have you lived here?`", 
                              "`9. What is your annual household income?`", 
                              "`10. Are you a student?`")

# And being able to loop through them means I can create stacked
# bar charts for every question with three lines. But it's
# also very computation-intensive so only uncomment it if
# you need it

#for (col in all_rating_questions) {make_stacked_bars(column = col, xticks = 11)}

```

### Who are the renters? Who are the owners?

One striking discovery is that renters and owners of homes differ noticeably in their responses to house prices (of course), night safety, and police trust. So who are the renters, and who are the owners?

``` {r own_v_rent}
for (col in all_rating_questions) {
  rent_own(column = col, xticks = 10)
}

for (col in all_categorical_questions) {
  rent_own(column = col, xticks = 10, categorical = TRUE)
}

renters_by_income = renters %>% dplyr::mutate(`9. What is your annual household income?` = factor(`9. What is your annual household income?`, levels=c("Less than $10,000", "$10,000 to $24,999", "$25,000 to $49,999", "$50,000 to 74,999", "$75,000 to $99,999", "$100,000 to $149,999", "$150,000 or more")))
renters_by_income %>% ggplot(aes_string(x="`9. What is your annual household income?`", fill="`6. Describe your housing status in Somerville.`")) + geom_bar(color = "black") + labs(x = "", y = "", title = "`9. What is your annual household income?`") + scale_fill_manual(values = c("steelblue1")) + theme(axis.text.x = element_text(angle = 75)) + theme(axis.text.x = element_text(vjust = 0.5)) 

owners_by_income = owners %>% dplyr::mutate(`9. What is your annual household income?` = factor(`9. What is your annual household income?`, levels=c("Less than $10,000", "$10,000 to $24,999", "$25,000 to $49,999", "$50,000 to 74,999", "$75,000 to $99,999", "$100,000 to $149,999", "$150,000 or more")))
owners_by_income %>% ggplot(aes_string(x="`9. What is your annual household income?`", fill="`6. Describe your housing status in Somerville.`")) + geom_bar(color = "black") + labs(x = "", y = "", title = "`9. What is your annual household income?`") + scale_fill_manual(values = c("seagreen1")) + theme(axis.text.x = element_text(angle = 75)) + theme(axis.text.x = element_text(vjust = 0.5)) 

```


### End chunk

This is just so I can click "run all the above" to run
the whole script at once.

``` {r wish_this_was_python}
# placeholder chunk
```
