---
title: "Data Manipulation Paper Trail"
author: "Peter Madsen and Chris Salas"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(readr)

somerville2017full <- read_csv("data/Happiness_Survey_2017.csv")
```

# Paper Trail

It turns out working with scripts in R is a huge pain, which furthers my impression that R is generally inferior to other languages in every way and only sees use because it's The Statistics Language(tm)*. I might do some of this in Python behind the scenes and hope people don't notice.

Anyway, the new plan is, each manipulation of the dataframe will sit in its own code chunk here with explanation. First, we'll separate out the demographics-related columns and do some histograms.

*: Let this language die, Math People

```{r demographics_dataframe}
somerville2017demographics <- somerville2017full[c(18:23, 25:27)]
gender_responses = unique(somerville2017demographics[colnames(somerville2017demographics)[1]])
# As can be seen in gender_responses, we have options beyond the simplistic gender binary (yay!) but also many redundant and/or ambiguous responses about multiple people filling out one survey which need to be cleaned (not yay).

#It's gonna take too long to figure out how to make R do this, so I'll clean up the data in Python and export it.

```

### Exercise 2

...

