---
title: "Data Calculations"
author: "Peter Madsen and Chris Salas"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(readr)
library(dplyr)
library(forcats)
library(ggplot2)

# The full data set with blanks removed.
fullsomerville2017 <- read_csv("cleanedData/2017HappinessSurvey.csv")

fullsomerville2017 = fullsomerville2017 %>% filter(`2. Age?` != "Blank" &
                                                     `9. What is your annual household income?` != "Blank" &
                                                     `4. What is your race or ethnicity?` != "Blank" &
                                                     `1. What is your gender` != "Blank" &
                                                     `7b. How would you rate the cost of housing?` != 0 &
                                                     `9. How safe do you feel walking in your neighborhood at night?` != 0 &
                                                     `7d. How would you rate your trust in the local police?` != 0)
fullsomerville2017 = fullsomerville2017 %>% filter(`4. What is your race or ethnicity?` == "White" | 
                                                     `4. What is your race or ethnicity?` == "Asian" | 
                                                     `4. What is your race or ethnicity?` == "Hispanic/Latino" | 
                                                     `4. What is your race or ethnicity?` == "Black/African American")

# Reference arrays so I know column names and such
the_columns = c("`Language submitted in:`",
                "`1. How happy do you feel right now?`",
                "`2. How satisfied are you with your life in general?`",
                "`3. How anxious did you feel yesterday?`",
                "`4. How satisfied are you with Somerville as a place to live?`",  # 5
                "`5. How satisfied are you with your neighborhood?`",
                "`6. How proud are you to be a Somerville resident?`",
                "`7a. How would you rate the availability of information about city services?`",
                "`7b. How would you rate the cost of housing?`",
                "`7c. How would you rate the overall quality of public schools?`",  # 10
                "`7d. How would you rate your trust in the local police?`",
                "`7e. How would you rate the maintenance of streets and sidewalks?`",
                "`7f. How would you rate the availability of social community events?`",
                "`8. How satisfied are you with the quality and number of transportation options available to you?`",
                "`9. How safe do you feel walking in your neighborhood at night?`",  # 15
                "`10. How satisfied are you with the appearance of parks and squares in your neighborhood?`",
                "`11. How satisfied are you with the beauty or physical setting of your neighborhood?`",
                "`1. What is your gender`",
                "`2. Age?`",
                "`3. What language(s) do you speak at home?`",  # 20
                "`4. What is your race or ethnicity?`",
                "`5. Do you have children age 18 or younger who live with you?`",
                "`6. Describe your housing status in Somerville.`",
                "`7. Do you plan to move away from Somerville in the next two years?`",
                "`8 How long have you lived here?`",  # 25
                "`9. What is your annual household income?`",
                "`10. Are you a student?`")

gender_subsets = c("Female", 
                     "Male", 
                     "Non-binary", 
                     "Both selected", 
                     "Multiple people")
age_subsets = c("18-24", 
                  "25-34", 
                  "35-44", 
                  "45-54", 
                  "55-64", 
                  "65-74", 
                  "75 or older")
income_subsets = c("Less than $10,000", 
                     "$10,000 to $24,999", 
                     "$25,000 to $49,999", 
                     "$50,000 to 74,999", 
                     "$75,000 to $99,999", 
                     "$100,000 to $149,999", 
                     "$150,000 or more")
ethnicity_subsets = c("White", 
                        "Asian", 
                        "Hispanic/Latino", 
                        "Black/African American")


# function definition for... getting important math.
get_important_math = function(column, dataframe = fullsomerville2017, categorical = FALSE) {
  if (categorical == TRUE) {
    for (i in column) {
      the_proportion = dataframe %>% count(dataframe[i]) %>% mutate(p = n / pull(count(dataframe)))
      print(the_proportion)
    }
  } else {
    for (i in column) {
      data_as_array = pull(dataframe, column)
      the_mean = mean(data_as_array)
      the_min = min(data_as_array)
      the_max = max(data_as_array)
      the_quartiles = quantile(data_as_array, probs = c(0.25,0.5,0.75))
      first_quartile = the_quartiles[1]
      the_median = the_quartiles[2]
      third_quartile = the_quartiles[3]
      the_standard_deviation = sd(data_as_array)
      the_message = paste("Mean: ", the_mean, ", Min: ", the_min, ", 1st Q: ", first_quartile, ", Median: ", the_median, ", 3rd Q: ", third_quartile, ", Max: ", the_max, ", StDev: ±", the_standard_deviation, sep = "")
      print(the_message)
    }
  }
}
```

# Calculations

## Goals

> 1. Build a function which calculates mean, standard deviation, and five number summary for quantitative variables or the proportion for categorical ones, given a dataframe and a column




```{r figure1}
# Figure 1 involves the six basic demographics: gender (18), age (19), income (26), student (27), ethnicity (21), housing status (23)
the_six_basic = c(18, 19, 26, 27, 21, 23)
get_important_math(the_six_basic, categorical = TRUE)


```

It is with blithe spirit that I receive this news, that the function works with categorical values. Now, we commence with quantitative ones.

``` {r figure2}
# Figure 2 is house price ratings (9) from owners and renters (23)
owners = fullsomerville2017 %>% filter(`6. Describe your housing status in Somerville.` == "Own")
renters = fullsomerville2017 %>% filter(`6. Describe your housing status in Somerville.` == "Rent")

get_important_math_for_OwnRent = function(column, own = owners, rent = renters) {
  print("For Owners,")
  get_important_math(column = column, dataframe = own)
  print("")
  print("For Renters,")
  get_important_math(column = column, dataframe = rent)
}

get_important_math_for_OwnRent(9)
```

Also yay, now it works for numbers, too!

``` {r figure3}
# house prices (9) by income (26)
income1=fullsomerville2017 %>% filter(`9. What is your annual household income?` == income_subsets[1])
income2=fullsomerville2017 %>% filter(`9. What is your annual household income?` == income_subsets[2])
income3=fullsomerville2017 %>% filter(`9. What is your annual household income?` == income_subsets[3])
income4=fullsomerville2017 %>% filter(`9. What is your annual household income?` == income_subsets[4])
income5=fullsomerville2017 %>% filter(`9. What is your annual household income?` == income_subsets[5])
income6=fullsomerville2017 %>% filter(`9. What is your annual household income?` == income_subsets[6])
income7=fullsomerville2017 %>% filter(`9. What is your annual household income?` == income_subsets[7])

get_important_math(9, income1)
get_important_math(9, income2)
get_important_math(9, income3)
get_important_math(9, income4)
get_important_math(9, income5)
get_important_math(9, income6)
get_important_math(9, income7)


```

``` {r figure4}
# house prices (9) by age (19)
age1=fullsomerville2017 %>% filter(`2. Age?` == age_subsets[1])
age2=fullsomerville2017 %>% filter(`2. Age?` == age_subsets[2])
age3=fullsomerville2017 %>% filter(`2. Age?` == age_subsets[3])
age4=fullsomerville2017 %>% filter(`2. Age?` == age_subsets[4])
age5=fullsomerville2017 %>% filter(`2. Age?` == age_subsets[5])
age6=fullsomerville2017 %>% filter(`2. Age?` == age_subsets[6])
age7=fullsomerville2017 %>% filter(`2. Age?` == age_subsets[7])

get_important_math(9, age1)
get_important_math(9, age2)
get_important_math(9, age3)
get_important_math(9, age4)
get_important_math(9, age5)
get_important_math(9, age6)
get_important_math(9, age7)

```

``` {r figure5}
# night safety (15) by gender (18)
the_men = fullsomerville2017 %>% filter(`1. What is your gender` == gender_subsets[2])
the_women = fullsomerville2017 %>% filter(`1. What is your gender` == gender_subsets[1])

get_important_math(15, the_women)
get_important_math(15, the_men)

```

``` {r figure6}
# Police trust (11) by income (26)
get_important_math(11, income1)
get_important_math(11, income2)
get_important_math(11, income3)
get_important_math(11, income4)
get_important_math(11, income5)
get_important_math(11, income6)
get_important_math(11, income7)


```

``` {r figure7}
# Police trust (11) by age (19)
get_important_math(11, age1)
get_important_math(11, age2)
get_important_math(11, age3)
get_important_math(11, age4)
get_important_math(11, age5)
get_important_math(11, age6)
get_important_math(11, age7)

```

``` {r figure8}
# Renters and Owners by income (26), categorical
get_important_math(26, dataframe = renters, categorical = TRUE)
get_important_math(26, dataframe = owners, categorical = TRUE)

```

``` {r figure9}
# Renters and Owners by age (19), categorical
get_important_math(19, dataframe = renters, categorical = TRUE)
get_important_math(19, dataframe = owners, categorical = TRUE)

```

``` {r figure10}
# Life in general (3) for renters and owners
get_important_math_for_OwnRent(3)
```

``` {r figure11}
# police trust (11) for renters and owners
get_important_math_for_OwnRent(11)
```

``` {r figure12}
# night safety (15) for renters and owners
get_important_math_for_OwnRent(15)
```

``` {r run_all}
# Chunk to run all above
```


